---
output: github_document
---

# The Checklist Package <img src="man/figures/logo.svg" align="right" alt="A hexagon with the word checklist" width="120" />
 
<!-- README.md is generated from README.Rmd. Please edit that file -->

[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental-1)
![GitHub](https://img.shields.io/github/license/inbo/checklist)
[![R build status](https://github.com/inbo/checklist/workflows/check%20package%20on%20main/badge.svg)](https://github.com/inbo/checklist/actions)
[![Codecov test coverage](https://codecov.io/gh/inbo/checklist/branch/master/graph/badge.svg)](https://codecov.io/gh/inbo/checklist?branch=master)
![GitHub code size in bytes](https://img.shields.io/github/languages/code-size/inbo/checklist.svg)
![GitHub repo size](https://img.shields.io/github/repo-size/inbo/checklist.svg)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.4028303.svg)](https://doi.org/10.5281/zenodo.4028303)

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

The goal of `checklist` is to provide an elaborate and strict set of checks for R packages and R code.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

```{r installation, eval = FALSE}
# install.packages("remotes")
remotes::install_github("inbo/checklist")
```

## Examples

You can run the full list of checks

```{r example-full, eval = FALSE}
library(checklist)
check_package() # for packages
check_source() # for a project with R and Rmd files
```

Or run the individual checks

```{r example-detail, eval = FALSE}
check_cran()
check_description()
check_documentation()
check_lintr()
check_filename()
```

Create a `checklist.yml` to allow some of warnings or notes.

```{r example-checklist, eval = FALSE}
write_checklist()
```

Start a package from scratch with everything set-up

```{r example-create, eval = FALSE}
create_package()
```
