---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Lifecycle:maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://lifecycle.r-lib.org/articles/stages.html#maturing-1)
![GitHub](https://img.shields.io/github/license/inbo/checklist)
[![License](https://img.shields.io/badge/license-GPL--3-blue.svg?style=flat)](https://www.gnu.org/licenses/gpl-3.0.html)
[![Release](https://img.shields.io/github/release/inbo/checklist.svg)](https://github.com/inbo/checklist/releases)
[![R build status](https://github.com/inbo/checklist/workflows/check%20package%20on%20main/badge.svg)](https://github.com/inbo/checklist/actions)
![r-universe name](https://inbo.r-universe.dev/badges/:name?color=c04384)
![r-universe package](https://inbo.r-universe.dev/badges/checklist)
[![Codecov test coverage](https://codecov.io/gh/inbo/checklist/branch/main/graph/badge.svg)](https://app.codecov.io/gh/inbo/checklist?branch=main)
![GitHub code size in bytes](https://img.shields.io/github/languages/code-size/inbo/checklist.svg)
![GitHub repo size](https://img.shields.io/github/repo-size/inbo/checklist.svg)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.4028303.svg)](https://doi.org/10.5281/zenodo.4028303)
<!-- badges: end --> 

# The Checklist Package <img src="man/figures/logo.svg" align="right" alt="A hexagon with the word checklist" width="120" />

**keywords**: quality control; documentation; publication
<!-- community: inbo -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.path = "man/figures/README-",
  out.width = "100%", eval = FALSE
)
```

The goal of `checklist` is to provide an elaborate and strict set of checks for R packages and R code.

## Installation

You can install the package from the [INBO universe](https://inbo.r-universe.dev/ui#builds) via

```{r universe}
# Enable the INBO universe
options(
  repos = c(
    inbo = "https://inbo.r-universe.dev", CRAN = "https://cloud.r-project.org"
  )
)
# Install the packages
install.packages("checklist")
```

If that doesn't work, you can install the version from [GitHub](https://github.com/inbo/checklist/) with:

```{r installation}
# install.packages("remotes")
remotes::install_github("inbo/checklist")
```

## Examples

You can run the full list of checks

```{r example-full}
library(checklist)
check_package() # for packages
check_source() # for a project with R and Rmd files
```

Or run the individual checks

```{r example-detail}
check_cran()
check_description()
check_documentation()
check_lintr()
check_filename()
```

Create a `checklist.yml` to allow some of warnings or notes.

```{r example-checklist}
write_checklist()
```

Start a package from scratch with everything set-up

```{r example-create}
create_package()
```
