---
title: "Organisations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Organisations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r setup, include = FALSE}
library(knitr)
opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## New philosophy

`checklist 0.5.0` had a major change in the way it handles organisation information.
The old `organisation` object was replaced by two new classes: `org_item` and `org_list`.
Another important change is that an organisation can define minimum rules for the organisation.
This set of rules is stored in a repository called `checklist` in the organisation's git organisation.

### Why bother to set an `org_list` object?

`check_package()` and `check_project()` use this information to check if the project is compliant with the organisation's rules about naming conventions, copyright holder and funder.
This allows the organisation to enforce a consistent set of rules across all its projects.

## The `org_item` class

The `org_item` class is used to store information about an organisation.
It contains the two mandatory fields: `name` and `email`.
The name must be a named vector using a language code as name.
This allows the organisation name to be used in different languages.
The first name will be used as the default name when no match is available.
The email served a double purpose:
1. Having an organisation email in the metadata allows users to contact the organisation in case that the individual authors are not available or do not respond.
2. `checklist` uses the domain of the email to match other persons to this organisation based on email.

Optionally, you can define the organisations [ROR identifier](https://ror.org/); if matching persons must having an [ORCID identifier](https://orcid.org/); if the projects must list a specific [Zenodo community](https://zenodo.org/communities#community-info) and what kind of licenses are allowed for packages, projects and data.

The `org_item` defines the required status of the organisation as a copyright holder or funder.
By default, this is set to `"optional"`, which means that the organisation is not required to be listed as a copyright holder or funder but it can be listed as such.
`"single"` is the other side of the spectrum, which means that the organisation must be listed as a copyright holder or funder and no other organisation can be listed as such.
`"shared"` is a weaker version of `"single"` where the organisation must be listed as a copyright holder or funder but other organisations can be listed as such too.
Finally, `"when no other"` means that the organisation must be listed as a copyright holder or funder when no other organisation is listed as such.

## The `org_list` class

The `org_list` class is used to store a list of `org_item` elements.
The rules of the different `org_item` within a single `org_list` must be compatible.
E.g. if one organisation is set to `"single"` as copyright holder, then no other organisation can be set to `"single"` or `"shared"` as copyright holder.

Additionally, it contains the URL of the main organisations git organisation.
The `checklist` repository at that URL is used to store the minimum rules for the organisation.
E.g. when the URL is set to `https://github.com/inbo`, then we look for the minimum rules at `https://github.com/inbo/checklist`.

## Using the organisations default `org_list`

### Setting the default

First you need to create the `org_item` objects that will make up the default `org_list`.

```{r create-org-item}
library(checklist)
inbo <- org_item$new(
  name = c(
    "nl-BE" = "Instituut voor Natuur- en Bosonderzoek (INBO)",
    "fr-FR" = "Institut de Recherche sur la Nature et les Forêts (INBO)",
    "en-GB" = "Research Institute for Nature and Forest (INBO)",
    "de-DE" = "Institut für Natur- und Waldforschung (INBO)"
  ),
  email = "info@inbo.be",
  ror = "https://ror.org/00j54wy13",
  orcid = TRUE,
  zenodo = "inbo",
  rightsholder = "shared",
  funder = "when no other",
  license = list(
    package = c("GPL-3 ", "MIT"), project = "CC BY 4.0", data = "CC0"
  )
)
inbo
anb <- org_item$new(
  name = c(
    `nl-BE` = "Agentschap voor Natuur en Bos (ANB)",
    `en-GB` = "Agency for Nature & Forests (ANB)"
  ),
  email = "natuurenbos@vlaanderen.be",
  ror = "https://ror.org/04wcznf70",
  license = list(
    package = character(0),
    project = character(0),
    data = character(0)
  )
)
anb
```

Then you can create the `org_list` object.

```{r create-org-list}
inbo_org <- org_list$new(inbo, anb, git = "https://github.com/inbo")
inbo_org
```

The final step is to store the `org_list` into a `checklist` repository.
The `write()` method creates a `organisation.yml` file in the root of the repository.
Commit this file to the repository and push it to the remote.

```{r store-org-list}
path_to_checklist_repo <- tempfile("checklist")
dir.create(path_to_checklist_repo, recursive = TRUE)
inbo_org$write(path_to_checklist_repo)
```

### Getting the default

Run `get_default_org_list()` in a project to get the default organisation list.
This function determines the organisation based on the git remote `origin` of the project.
Hence it only works for git projects.
It look for the an `organisation.yml` file in the root of the `main` branch of the repository `checklist` of the git organisation.
Note that you can have projects from different organisations on the same machine.

`get_default_org_list()` stores the information in the configuration settings of `checklist`.
You only need to run it once for every organisation in case the organisations defaults change.
After that you can simply run `org_list$new()$read()`.
That function either returns the local `organisation.yml` file or fetches the default from the `checklist` configuration.

```{r get-org-list}
# setting up a demo git repository
test_dir <- tempfile("test")
dir.create(test_dir)
library(gert)
git_init(test_dir)
git_remote_add(
  url = "https://github.com/ThierryO/test", name = "origin", repo = test_dir
)
# get the default organisation list from https://github.com/ThierryO/checklist
my_org <- get_default_org_list(test_dir)
my_org
```

### Adding an `org_item` to an `org_list` object

You can add an `org_item` to an `org_list` object using the `add_item()` method.
At this point it is required to use the `write()` method to store the `org_list` object in the local `organisation.yml` file.

```{r update-org-list}
# let's create an empty org_list so we can use its methods
empty_org <- org_list$new()
# read the organisation list
# because we didn't write the organisation.yml file, it will fetch the
# default from the configuration
list.files(test_dir)
empty_org$read(test_dir)
# add an organisation item to the organisation list
my_org <- my_org$add_item(anb)
my_org
# it still isn't written to the local organisation.yml file
list.files(test_dir)
empty_org$read(test_dir)
# use the write method to create the organisation.yml file
my_org$write(test_dir)
list.files(test_dir)
empty_org$read(test_dir)
```

## FAQ

1. Can I use `checklist` without git?

   Yes, you can use `checklist` without git.
   `checklist` will only check the rules defined in the local `organisation.yml`

1. Is a `checklist` repository at the organisation level required to use `checklist`?

   No, you can use `checklist` without a `checklist` repository.
   Having no `checklist` repository means that you don't enforce any organisation rules.


1. Do I need to add an `organisation.yml` file to my project?

   No, you do not need to add an `organisation.yml` file to your project unless you want to add specific rules for your project.
   Note that these still need to comply with the organisation rules defined in the `checklist` repository.
