---
title: "Spell checking with checklist"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Spell checking with checklist}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(knitr)
opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Configuration

The configuration of `check_spelling()` is stored in the `checklist.yml` file at the root of the project.
This comprises the default language to use for spell checking and optionally which files to exclude from spell checking or to spell check using a different language.
Use `setup_package()` or `setup_project()` first.
We recommend to follow the steps in `vignette("getting_started", package = "checklist")` to set-up the `checklist.yml`.
In case of a package you must set the language in the `Language:` field of the `DESCRIPTION` file.
In case the `Language:` field is not set, `en-GB` is used.
`setup_project()` will ask which default language to set and will store the information in `checklist.yml`.

Note that a language is defined with the `xx-YY` syntax.
The first two letters must be lower case (`xx`) and define the language.
The last two letters must be upper case (`YY`) and define the language variant.
A few relevant examples:

- `en-GB`: English as written in the UK.
- `en-US`: English as written in the USA.
- `nl-BE`: Dutch as written in Belgium.
- `nl-NL`: Dutch as written in the Netherlands.

Before you can change the setting, you must read the current settings with `x <- read_checklist(path)`.
This object has several methods to change the spell checking settings.
`x$set_default(language)` changes the default language.
`x$set_exceptions()` allows you to review the current settings on the relevant folders and files within the project.
You must choose between keeping the current configuration, using the default language for all files or change the setting for some files.
If you select the last options, the function starts at the root of the project and lists the relevant files or folders at the root.
In case of a file, you can choose to ignore the file, use the default language or use an additional language.
If you select any of these options in case of a folder, it will recursively apply to all files in that folder.
To change the setting for some of the files, the function gives an extra option in case of a folder.
After updating the settings, you need to store them with `write_checklist(x)`.
`check_spelling()` will now use the updated settings.

```{r eval = FALSE}
# This code assumes your working directory is at the root of your project.
# Pass the path to the project root to read_checklist() if that is not the case.
x <- read_checklist()
x$set_default("nl-BE")
x$set_exceptions()
write_checklist(x)
```

## Basic usage

Run `check_spelling()` at the working directory of the project.
Or run `check_spelling(path)` where `path` points to a directory with the files you want to spell check.
This will look for all markdown files (`.md` and `.Rmd`) within the directory and its subdirectories.
In case your project is an R package, it will also check all R help files (`.Rd`) in the `man` directory.
Note that `check_spelling()` does not check R code.
Neither in `.R` files nor chunks in `.Rmd` files.

The function returns a data.frame with every word not present in the dictionary.
Displaying the data.frame in an R session outside RStudio lists all unknown words per file combined with the line and column number were they appear.
When you display this data.frame in RStudio, it opens a `Markers` tab.
Clicking on an issue in this tab will open the corresponding file and move the cursor to that line.

Some of the issues will be typos and need to be fixed.
Others are words not listed in the available dictionaries.
The following options work both with `.Rmd` as `.Rd` files.

1. `check_spelling()` ignores words with `verbatim` syntax (between back ticks "`").
1. Exclude the entire file from spell checking.
More details on that in the configuration section.
1. Add the words to a project specific custom dictionary (see below).

You have a few more options to handle those words in `.Rmd` files.

1. You can ignore an entire line by placing `<!-- spell-check: ignore -->` at the end of the line.
1. Ignore a block of lines by adding an extra line with `<!-- spell-check: ignore:start -->` at the beginning of the block.
End the block with `<!-- spell-check: ignore:end -->` on a new line.

## Custom dictionary

First of all make sure that `check_spelling()` only returns words you want to add to the custom dictionary.
Next store the output of `check_spelling()` in an object and use that object as input in `custom_directory()`.
This will append to words to a `.dic` file in the `inst` folder.
Running `check_spelling()` again should return an empty list of issues.

```{r, eval = FALSE}
issues <- check_spelling()
custom_dictionary(issues)
```

> I use a few technical terms quit often.
> Can I use a common custom dictionary instead of a project based custom dictionary?

No.
Project based custom dictionaries are located within the project.
This is required to make the project reproducible.
What you can do is copy the `.dic` files from other projects.

# Quarto projects

When `check_spelling()` finds a [Quarto book project](https://quarto.org/docs/books/), it will check if `lang` is set in `_quarto.yml`.
Then `check_spelling()` uses that `lang` for all files listed under the `book: chapters` and `book: appendices` tabs.
Please use the Quarto specific markdown to specify sections with a different language.
Here are two examples.

```
    ::: {lang=nl-BE}
    Paragraph in Dutch
    :::
```

```
    A sentence with a [word]{lang=nl-BE} in Dutch.
```
